---
title: "Sensitivity Analysis Almond Yield"
author: "AnnaClaire Marley"
date: "4/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Reading in libraries and dataset
```{r, include = FALSE}
# load packages
library(tidyverse)
library(pse)
library(sensitivity)
library(gridExtra)

# source function
source('almond_anomaly.R')
```

## Sensitivity Analysis on Almond Yield Function

* **Perform a sensitivity analysis of how mean anomaly varies all of the parameters used in the almond yield model**
* **Assume parameters are normally distributed with standard deviation of 20% mean value**

#### **Latin Hypercube Method**

```{r}
# All parameters of almond yield model
factors = c("a", "b", "c", "d", "e")

# Decide How many parameter sets to run
nsets = 1000

# choose distributions for parameters - 
q = c("qnorm", "qnorm", "qnorm", "qnorm", "qnorm")
q.arg = list(list(mean = -0.015, sd = 0.015/5), 
             list(mean = -0.0046, sd = 0.0046/5), 
             list(mean = -0.07, sd = 0.07/5),
             list(mean = 0.0043, sd = 0.0043/5), 
             list(mean = 0.28, sd = 0.28/5))

# generate samples from LHS
sens_almond = LHS(NULL, factors, nsets, q, q.arg)

# set of values generaated for each parameter
sens_pars = get.data(sens_almond)
head(sens_pars)

# now run the model for each set of parameter samples
# first create a data structure to store results
sens_results = matrix(nrow=nsets, ncol=1)

# read in the input data
clim <- read.table("clim.txt")

# run the almond yield model for each parameter set
tmp = mapply(FUN=almond_anomaly, 
             a = sens_pars$a, 
             b = sens_pars$b, 
             c = sens_pars$c, 
             d = sens_pars$d,
             e = sens_pars$e,
             MoreArgs=list(clim_data = clim,
                           mean_only = TRUE))
#head(tmp)

# turn into matrix using unlist 
sens_results_LH = matrix((unlist(tmp)), ncol=1, byrow=TRUE)
colnames(sens_results_LH) = c("mean_yield")


```

**Rank the parameters in term of their sensitivity**

```{r}

# send results back
# tell() provides the sensitivity structure created by LHS, results from model simulations
sens_almond = pse::tell(sens_almond, t(sens_results_LH), 
                        res.names=c("mean_yield"))
# plot sensitivity
pse::plotscatter(sens_almond, col="blue", cex=5)

# Plot the partial rank correlation coefficient
pse::plotprcc(sens_almond)
sens_almond$prcc

```

Sensitivity ranking

* 1) d (precip^2 coefficient) 
* 2) c (precip coefficient)
* 3) b (temp^2 coefficient) 
* 4) e (model intercept)
* 5) a (temp coefficient)


**Graph uncertainty in mean yield anomaly across all parameter uncertainty (boxplot and cummulative distribution of the output).**

```{r}
# make it a dataframe
sens_results_LH = as.data.frame(sens_results_LH)

# boxplot
boxplot_LH <-  ggplot(data = sens_results_LH, aes(x = " ", y = mean_yield)) +
                  geom_boxplot(fill = "brown") +
                  labs( 
                   title = "Latin Hypercube Uncertainty",
                    x = "Almond",
                    y = "Mean Yield Anomaly (ton/acre)"
                    ) +
                  theme_classic()
  boxplot_LH
# cumulative distribution   
sens_almond_cdf = pse::tell(sens_almond, sens_results_LH, res.names=c("mean_yield"))

plotecdf(sens_almond_cdf, col = "red", lwd = 5, 
         xlab = "Mean Almond Yield Anomaly",
         main = "Empirical Cumulative Distribution Latin Hypercube Method")

# or using ggplot 
cum_dist_lh_gg <- ggplot(sens_results_LH, aes(mean_yield)) +
                  stat_ecdf(geom = "step") +
                  xlab("Mean Yield Anomaly (ton/acre)") +
                  ylab("Proportion") +
                  labs(title = "Cumulative Distribution LHS Method") +
                 theme_classic()
cum_dist_lh_gg
```


#### **Sobel method**

Workflow:

* run sobel to get parameter sets in a sensitivity analysis object
* run model with those parameter sets
* tell the senstivity object about results associated with each parameter set
* look at sensitivity analysis metric from sobel

```{r}

# number of runs
np=1000

### generate two examples of random number from parmeter distributions ####

  # naming parameters
  a = rnorm(mean = -0.015, sd = 0.015/5, n=np)
  b = rnorm(mean = -0.0046, sd = 0.0046/5, n=np)
  c = rnorm(mean = -0.07, sd = 0.07/5, n=np)
  d = rnorm(mean = 0.0043, sd = 0.0043/5, n=np)
  e = rnorm(mean = 0.28, sd = 0.28/5, n=np)
  
  # create dataframe
  X1 = cbind.data.frame(a, b, c, d, e)
  
  ## repeat process for second dataframe
  # naming parameters
  a = rnorm(mean = -0.015, sd = 0.015/5, n=np)
  b = rnorm(mean = -0.0046, sd = 0.0046/5, n=np)
  c = rnorm(mean = -0.07, sd = 0.07/5, n=np)
  d = rnorm(mean = 0.0043, sd = 0.0043/5, n=np)
  e = rnorm(mean = 0.28, sd = 0.28/5, n=np)
  
  # create dataframe
  X2 = cbind.data.frame(a, b, c, d, e)

# run sobel sensitivity
sens_sobel = sobol2007(model = NULL, X1, X2, nboot = 100)

# run model with parameters generated from sobel
res = mapply(FUN = almond_anomaly,
             a = sens_sobel$X$a,
             b = sens_sobel$X$b,
             c = sens_sobel$X$c,
             d = sens_sobel$X$d,
             e = sens_sobel$X$e,
             MoreArgs=list(clim_data = clim,
                           mean_only = TRUE))


sens_sobel_tell = sensitivity::tell(sens_sobel, res, res.names="mean_yield")

print(sens_sobel_tell)
plot(sens_sobel_tell)


```

```{r}
#For testing purpose
sum(sens_sobel_tell$S$original)
sum(sens_sobel_tell$T$original)
```


**Rank the parameters in term of their sensitivity**

Sensitivity ranking (First Order)

* 1) d (precip^2 coefficient) 
* 2) c (precip coefficient)
* 3) e (model intercept) 
* 4) a (temp coefficient) 
* 5) b (temp^2 coefficient)

Sensitivity ranking (Total)

* 1) d (precip^2 coefficient) 
* 2) c (precip coefficient)
* 3) e (model intercept) 
* 4) a (temp coefficient) 
* 5) b (temp^2 coefficient)


**Graph uncertainty in mean yield anomaly across all parameter uncertainty (boxplot and cummulative distribution of the output).**

```{r}

# turn into matrix using unlist 
results_sobel = matrix((unlist(res)), ncol=1, byrow=TRUE)
colnames(results_sobel) = c("mean_yield")

# make a data frame for plotting
results_sobel_df <- as.data.frame(results_sobel)

# boxplot
boxplot_sobel <- ggplot(data = results_sobel_df, aes(x = " ", y = mean_yield)) +
          geom_boxplot(fill = "green") +
          labs( 
           title = "Sobel Uncertainty",
            x = "Almond",
            y = "Mean Yield Anomaly (ton/acre)"
            ) +
          theme_classic()

boxplot_sobel

# cumulative distribution   
cum_dis_sobel <- ggplot(results_sobel_df, aes(mean_yield)) +
                stat_ecdf(geom = "step") +
                xlab("Mean Yield Anomaly (ton/acre)") +
                ylab("Proportion") +
                labs(title = "Cumulative Distribution Sobel Method") +
                theme_classic()
cum_dis_sobel
```

#### All graphs

```{r}

grid.arrange(boxplot_LH, cum_dist_lh_gg, boxplot_sobel, cum_dis_sobel, ncol = 2)

```

