---
title: "Crop Yields"
author: "AnnaClaire Marley"
date: "4/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}
# load packages
library(tidyverse)
```

```{r}
# read in the data
clim <- read.table("../assignment_2/clim.txt")
```

**1. Implement a model of almond yield anomaly (difference from average) in R based on this paper; You will be graded both on the “correctness” of the implementation and on “style”. Make sure you add some error checking. Use coefficients in the paper as default values, but make sure you can vary them. **

**2. Have the function return almond yield anomaly for each year, and max and minimum yields over a time series of multiple year inputs**

*Almond model from Lobell et al:*

Y = -0.015T<sub>n,2</sub> - 0.0046T<sup>2</sup><sub>n,2</sub> - 0.07P<sub>1</sub> + 0.0043P<sup>2</sup><sub>1</sub> + 0.28

- Y: yield anomaly (ton/acre)
- Tn: min temperature (degrees C)
- Tx: maximum temperature
- P: precipitation (mm)
- subscripted numbers represent month of climate variable.
  - So almond yields need min temp from feb and precipitation from jan
  

```{r function}

calc_crop_yield = function(df = clim, a_param = -0.015, b_param = -0.0046, c_param = -0.07, d_param = 0.0043, intercept = 0.28) {
  
  # new dataframe of mean monthly min temp and monthly precipitation
  clim_month = clim %>% 
  group_by(year, month) %>% 
  summarize(meantmin = mean(tmin_c),
            precip=sum(precip)) %>% 
  ungroup()
  
  # make an empty dataframe to put results in
  number_years = length(unique(clim_month$year))
  crop = as.data.frame(matrix(nrow=number_years, ncol=2))
  colnames(crop)=c("year","yield_anom")
 
   for (i in 1:length(unique(clim_month$year))){
    success = TRUE
    # fill in the year column 
    crop$year[i] = unique(clim_month$year)[i]
    
    # choose specific month each year to extract min temp
    tmin_df <- clim_month %>% 
      filter(year == unique(clim_month$year)[i]) %>% 
      filter(month == 2) %>% 
      select(meantmin) 
    
      # error checking that there is data for tmin for that year
      if (nrow(tmin_df) != 1){
        cat("Warning: Missing data for tmin in", unique(clim_month$year)[i], "\n")
        success = FALSE
      }
    
    #choose specific month each year to precipitation
    precip_df <- clim_month %>% 
      filter(year == unique(clim_month$year)[i]) %>% 
      filter(month == 1) %>% 
      select(precip) 
    
      #error checking that there is data for precip for that year
      if(nrow(precip_df) !=1){
        cat("Warning: Missing data for precip in", unique(clim_month$year)[i], "\n")
        success = FALSE
      } else if(precip_df[[1]] < 0 ){ #error checking that precipitation is >= to 0
        cat("Warning: Precip is less than 0 in", unique(clim_month$year)[i], "\n")
        success = FALSE
      }
    
    # calculate yield anomaly
    if (success == TRUE) {
      crop$yield_anom[i] = a_param*tmin_df[[1]] +
                            b_param*tmin_df[[1]]^2 + 
                            c_param*precip_df[[1]] +
                            d_param*precip_df[[1]]^2 + 
                            intercept
    }
  }
  
  max = crop %>% 
    filter(yield_anom != "NA") %>% 
    summarize(max_yield = max(yield_anom))
  
  min = crop %>% 
    filter(yield_anom != "NA") %>% 
    summarize(min_yield = min(yield_anom))
  
  return(list(data = crop, max = max[[1]], min = min[[1]]))
  
}

```


```{r}

calc_crop_yield(clim)

```




**3. With the climate data provided (clim.txt), *use your model* to estimate how almond yield anomalies for each year in the record - graph the results.**



**4. Use model to estimate  the mean annual yield anomaly if the last 20 years had been 2C warmer**



**5. Vary one of the parameter by sampling from a normal distribution with mean the value from Lobell et al., (2006) and standard deviation 10% of the mean -  create a box plot of  mean annual yield anomaly for baseline and 2C climate (where boxes show variation due to uncertanty in the parameter) **


