---
title: "Forest Growth Model"
author: "AnnaClaire Marley"
date: "5/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(tidyverse)
library(deSolve)
```


###**Model of forest growth (where forest size in measured in units of carbon (C)):**

* dC/dt = 0 when air temperatures are below 0

* dC/dt  = r*C for forests where carbon is below a canopy closure threshold 

* dC/dt = g. for forests where carbon is above the threshold canopy closure

* dC/dt = 0 once carrying capacity is reached.

*The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon* 

**Canopy closure threshold:** the size of the forest at which growth rates change from exponential to linear

**r**: an early exponential growth rate 

**g**: the linear growth rate 



#### Run the model for 300 years (using the ODE solver)  starting with an initial forest size of 10 kg/C, and using the following parameters:

* canopy closure threshold (canopy_close) of 100 kgC 

* a carrying capacity (k) of 500 kg C

* a early growth rate (r) of 0.01

* a later growth rate (g) of 2 kg/year

* an air temperature (temp) of 20C

```{r}
#### parameters ###
years = seq(from=1, to=300, by=2)
initial_forest = 10 # (kg C)
canopy_close = 100 # (kg C)
carry_capacity = 500 # (kg C)
initial_growth = 0.01 # (kg/year)
linear_growth = 2 # (kg/year)
temp = 20 # degrees C

# create parameter list
parms = list(canopy_thresh = canopy_close, K = carry_capacity,
             r = initial_growth, g = linear_growth, temp = temp)

# source model
source('../assignment_4/forest_growth.R')

# run ODE model
results = ode(y=initial_forest, times=years, func=forest_growth, parms=parms)
colnames(results)=c("year","forest_carbon")
head(results)

```


##### Graph the results

```{r}
ggplot(as.data.frame(results), aes(x = year, y = forest_carbon)) +
  geom_line() +
  labs(
    x = "Years",
    y = "Forest Size (kg C)"
  ) +
  theme_classic()

```


##### Run a sobol sensitivity analysis that explores how the estimated maximum forest size varies with early and later growth rates. 

##### Graph the results of the sensitivity analysis as a box plot and a plot of the two sobol indices (S and T)


```{r}

# rate of forest growth 
  carb_change = parms$r*C
  
  # forest growth is 0 when temperatures are below 0
  carb_change = ifelse(parms$temp < 0, 0, carb_change)
  
  # forest growth is 0 when carrying capacity is reached 
  carb_change = ifelse(carb_change >= parms$K, 0, carb_change)
  
  # forest growth becomes linear when carbon is above the threshold canopy closure
  carb_change = ifelse(carb_change > parms$canopy_thresh, parms$g, carb_change)


```


